Horizontal and vertical diversity drive total biomass and its temporal stability in fish communities

### 2020-05-28 manuscript/bef_stability/manuscript.Rmd
Fix result PCA with two axis

```{r, echo = FALSE, results = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

@@

```{r}
knitr::opts_chunk$set(
  results = TRUE 
)
```

@@

Biodiversity is essential to determine and maintain ecosystem properties
[@duffy_why_2009].  Current biodiversity losses and global changes question the
way in which ecosystems will be affected in turn [@bellard_impacts_2012]. Since
1970s and the @may_will_1972 seminal paper, ecology aims to uncover the complex
relationships between biodiversity and ecosystem properties [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018]. Because biodiversity largely
determines both quantity and distribution of biomass or energy across
ecosystems, it has a tremendous effects on the properties exhibited by an
ecosystem, such as its total biomass [@hector_plant_1999;
@tilman_biodiversity_2014; @grace_integrative_2016;
@pennekamp_biodiversity_2018]], its biomass production
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016]
and the temporal stability of its biomass [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006;
@olivier_independent_nodate].

@@

Biodiversity can be quantified in numerous ways, as the number of species, i.e.
species richness and in number of functional groups, i.e species are
gathered on what they do. Those metrics can be gathered into "horizontal
diversity", describing the list of components present in a community, i.e. here
species and functional groups. An
orthogonal view, based on trophic chains, described species on what they eat. It 
can be defined as interactions (or biomass/energy fluxes) between species and how
they are distributed among communities. It can be quantified as the number
of interaction between species or the number of trophic level. This orthogonal
view of seeing biodiversity, can be gathered into "vertical diversity".  
Horizontal and vertical diversity are two views of quantifying biodiversity that
are, however, not independent from each other [@duffy_functional_2007; @zhao_horizontal_2019]. The add
or removal of one species can have different effects depending on its vertical
position in the trophic chain [@duffy_functional_2007]. Hence, it is crucial to
study their interplay to improve our understanding of ecosystem dynamics. 
(*This paragraph is surely too simplistic*)

@@

Environment also largely structure communities and is important to take in
account when considering natural communities. Environment directly affects
organism performance, and subsequently, abundance of species and in turn
determines total biomass [Add general ref; @grace_integrative_2016] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @friedpetersen_drivers_2020]. Hydrology of streams
constraints the composition of fish communities [REF toulouse à balle], and is
also related to fish population temporal stability [@franssen_annual_2011].
However it is not clear which is the relative strength of direct and indirect
effects of environment on total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

@@

We used an standardized monitoring of fish communities in french streams. 99
sites were followed more than 10 years on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow, enrichment and
their temporal variability (expect for altitude and slope). Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

@@

```{r}
myload(nmds, habitat_pressure, dir = mypath("report"))
st_basin <- get_basin_station(sf_obj = FALSE)
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)

# Compute stability, network metrics, etc with ;he new datasets
##myload(network_metrics, dir = mypath("data", "classes"))
#network_metrics %>%
  #left_join(op_analysis, by = "opcod") %>%
  #select(opcod, station, composition) %>%
  #unnest(composition) %>%
  #group_by(station, troph_group) %>%
  #summarise(richness_tot = length(unique(species)))
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb,
 type_network_metrics = "classes" 
)
# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync_std"]],
  nmds = NULL,
  basin = st_basin
  )

@@

```{r, eval = FALSE}
stab_div_troph <- com_data$tps_bm_troph %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_rich_std = log10(rich_std_med),
    log_rich_tot = log10(richness_tot),
    log_rich_tot_std = log10(rich_tot_std),
    log_bm = log10(biomass_med),
    log_bm_std = log10(bm_std_med),
    log_stab = log10(biomass_stab),
    log_stab_std = log10(bm_std_stab)
    ) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, richness_tot, log_rich, log_rich_std, log_rich_tot, log_rich_tot_std, log_stab, log_stab_std, log_bm, log_bm_std, biomass_med, bm_std_med, richness_tot, rich_tot_std) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

```{r}
myload(
  hab_analysis,
  temporal_station_desc,
  temporal_press_polluants,
  geo_station,
  dir = mypath("data")
)
habitat_press <- build_habitat_pressure_dataset(
  .habitat_analysis = hab_analysis,
  .tmp_press = temporal_press_polluants,
  .tmp_st_desc = temporal_station_desc,
  .geo_st = geo_station 
)
```

```{r, fig.show = "hide"}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler", "DBO_med")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv", "DBO_cv")
pca_pressure <- habitat_press[, names(habitat_press) %in% c(evt_var_mean, evt_var_cv, "station")] %>%
  na.omit

pca_rotated <- compute_rotated_pca(
  .data = pca_pressure[, !names(pca_pressure) %in% "station"],
  naxis = 5 
)

prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))

tot_var_expl_pca <- round(sum(pca_rotated$rotated$Vaccounted["Proportion Var", ]) * 100) 
```

@@

For each fish station, environmental variables were matched with the year of
fish sampling (except for water temperature for which the record began in 2006).
Then, the average temperature, flow, BOD, river width and depth were computed as
well as coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and
environmental variables are often collinear, a principal component analysis (PCA)
was performed on the scaled variables. The PCA explained `r tot_var_expl_pca`%
of the variance. An axis rotation was performed to maximize the representation
of the variables by the principal components (Fig. \@ref(fig:pca-rotated)).
Five principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature, (3) average enrichment (BOD) and CV of flow, (4) river size
CV (CV of river width and depth) and (5) CV of enrichment (CV of BOD). Further
informations about data transformation and filtering are detailed in
supplementary information.

@@

Structural Equation Modeling are able to assess multivariate relationships,
directs and indirects between variables.  The metamodel describes how variables
are interrelated (Figure \@ref(fig:matmet) A). We built separated structural
equation model for total biomass and its temporal variability (Figure
\@ref(fig:sem-plot) A & B). Structural equation modeling was set up according to the
hypothesis described in introduction and detailed in supplementary information.
Structural equation models were based on Gaussian linear models, some variables
were logged to fulfill model assumptions (see residual plots and data
transformation in supplementary
informations). To account for heterogeneity of hydrographic basin, we set it as random effect on the intercept. Moreover, statistical model were checked for the presence of
multicollinearity (Supplementary information). The slopes reported in the main
text are standardized in order to compare their magnitude ($r_\delta = \beta
\frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope, $\sigma_x$
and $\sigma_y$ being respectively the standard deviation of the predictive
variable $x$ and of the response variable $y$). In the main text, only the
slopes which had $P$ value inferior to $0.05$ were reported (but see Table SXX).
Linear models and SEMs were computed with respectively the `nlme` and
`piecewieseSEM` R packages. The code used for the analysis are accessible on
github.

@@

@@

```{r mod-div-total, eval = FALSE}
stab_div_mod <- lmer(
  data = sem_data,
  control = ctrl_lmer,
  formula = log_stab_std ~ log_rich_tot_std + (1 | basin)
)
stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
stab_div_boot <- bootMer(stab_div_mod, mySumm2, nsim = 100)
stab_div_ci <- confint(stab_div_boot)
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred <- ggpredict(stab_div_mod, terms = c("log_rich_tot_std"), type = "fe")
stab_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_stab_std = predicted,
    log_rich_tot_std = x,
    basin = group
  )
stab_div_pred %<>%
  mutate(
    stab = 10^log_stab_std, # FALSE
    rich = 10^log_rich_tot_std,
    cl10 =  10^conf.low,
    ch10 =  10^conf.high
  )

#
bm_div_mod <- lmer(data = sem_data,
  formula = log_bm_std ~ log_rich_tot_std + (1 | basin)
)
bm_div_boot <- bootMer(bm_div_mod, mySumm2, nsim = 100)
bm_div_ci <- confint(bm_div_boot)
bm_div_sum <- piecewiseSEM::rsquared(bm_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred <- ggpredict(bm_div_mod, terms = c("log_rich_tot_std"), type = "fe")
#plot(pred, add.data = TRUE)
bm_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_bm_std = predicted,
    log_rich_tot_std = x,
    basin = group
  )

# get coeff
coef_div <- map(list(stab = stab_div_mod, bm = bm_div_mod), fixef)
```

```{r mod-div-troph, fig.show = "hide", eval = FALSE}
stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab_std ~ log_rich_tot_std + troph_group + log_rich_tot_std:troph_group + (1 | basin)
  #control = ctrl
)

stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab_std ~  troph_group + (1 | basin)
  #control = ctrl
)

stab_troph_boot <- bootMer(stab_troph_mod, mySumm2, nsim = 100)
stab_troph_ci <- confint(stab_troph_boot)

#stab_troph_colin <- car::vif(stab_troph_mod)
stab_troph_sum <- piecewiseSEM::rsquared(stab_troph_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

#pred_troph_mod <- ggpredict(stab_troph_mod, terms = c("log_rich_tot_std", "troph_group"), type = "fe")
##plot(pred_troph_mod, add.data = TRUE)
#stab_troph_pred <- pred_troph_mod %>%
  #as_tibble() %>%
  #rename(
    #log_stab_std = predicted,
    #log_rich_tot_std = x,
    #troph_group = group
  #)

## Productivity and troph group

#bm_troph_mod <- lmer(data = na.omit(stab_div_troph),
  ##fixed = log_bm ~ log_rich_tot * troph_group,
  #formula = log_bm_std ~ log_rich_tot_std + troph_group + log_rich_tot_std:troph_group + (1 | basin)
  ##random = ~ 1 + log_rich_tot * troph_group | basin,
  ##control = ctrl 
#)
#bm_troph_boot <- bootMer(bm_troph_mod, mySumm2, nsim = 100)
#bm_troph_ci <- confint(bm_troph_boot)

#bm_troph_sum <- piecewiseSEM::rsquared(bm_troph_mod, method = "nagelkerke") %>%
  #mutate_at(vars(Conditional, Marginal), ~round(., 2))

#pred_troph_mod <- ggpredict(bm_troph_mod, terms = c("log_rich_tot_std", "troph_group"), type = "fe")
##plot(pred_troph_mod, add.data = TRUE)
#bm_troph_pred <- pred_troph_mod %>%
  #as_tibble() %>%
  #rename(
    #log_bm_std = predicted,
    #log_rich_tot_std = x,
    #troph_group = group
  #)

## get coeff
#coef_troph <- map(list(stab = stab_troph_mod, bm = bm_troph_mod), fixed.effects)
```
```{r stab-div-troph, eval = FALSE}
stab_div_by_troph <- compute_stat_troph_rich(
  .data = stab_div_troph,
  resp = "log_stab_std"
)

bm_div_by_troph <- compute_stat_troph_rich(
  .data = stab_div_troph,
  resp = "log_bm_std"
)
troph_perm_test <- map(c(stab = "log_stab_std", bm = "log_bm_std"),
  ~test_troph(.data = stab_div_troph, var = .x)
  )
troph_perm_test$bm$obs
troph_perm_test$stab$obs
troph_perm_test$bm$p_val
troph_perm_test$stab$p_val
```

@@

def_env_x_pos <- c(from = 0, to = 6)

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_stab_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*30/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2"),
  metamodel = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "stab_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*30/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2")
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_bm_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*45/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2"),
  metamodel = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "bm_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*45/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2")
)
```

@@

```{r stab-network, fig.show = "hide"}
sem_data %>%
  select(sync, cv_sp, ct, t_lvl) %>%
  gather(network, metrics, ct, t_lvl) %>%
  gather(stab_comp, value, sync, cv_sp) %>%
  ggplot(aes(y = value, x = metrics)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(stab_comp ~ network, scales = "free_x")

stab_ct_mod <- lmer(data = sem_data,
  #fixed = log_bm ~ log_rich_tot * troph_group,
  formula = log_stab_std ~ ct + (1 | basin)
  #random = ~ 1 + log_rich_tot * troph_group | basin,
  #control = ctrl 
)
stab_ct_boot <- bootMer(stab_ct_mod, mySumm2, nsim = 100)
stab_ct_ci <- confint(stab_ct_boot)

stab_ct_sum <- piecewiseSEM::rsquared(stab_ct_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred_stab_ct_mod <- ggpredict(stab_ct_mod, terms = c("ct"), type = "fe")
#plot(pred_troph_mod, add.data = TRUE)
stab_ct_pred <- pred_stab_ct_mod %>%
  as_tibble() %>%
  rename(
    log_stab_std = predicted,
    log_rich_tot_std = x,
    troph_group = group
  )

# get coeff
coef_stab_ct <- map(list(stab = stab_ct_mod), fixed.effects)

stab_tlvl_mod <- lmer(data = sem_data,
  #fixed = log_bm ~ log_rich_tot * troph_group,
  formula = log_stab_std ~ t_lvl + (1 | basin)
  #random = ~ 1 + log_rich_tot * troph_group | basin,
  #control = ctrl 
)
stab_tlvl_boot <- bootMer(stab_tlvl_mod, mySumm2, nsim = 100)
stab_tlvl_ci <- confint(stab_tlvl_boot)

stab_tlvl_sum <- piecewiseSEM::rsquared(stab_tlvl_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred_stab_tlvl_mod <- ggpredict(stab_tlvl_mod, terms = c("t_lvl"), type = "fe")
#plot(pred_troph_mod, add.data = TRUE)
stab_tlvl_pred <- pred_stab_tlvl_mod %>%
  as_tibble() %>%
  rename(
    log_stab_std = predicted,
    log_rich_tot_std = x,
    troph_group = group
  )

# get coeff
coef_stab_tlvl <- map(list(stab = stab_tlvl_mod), fixed.effects)
```

@@

rich_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_rich_tot_std")
tlvl_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "t_lvl")
RC1_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC1")
RC2_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC2")
```

@@

caption_indir_bm_sem <- paste0(
  "Indirect and total effect of variables on total biomass from the
  structural equation model. See the table 1 for details about computation and
  meaning."

@@

```{r indir-stab}
format_table <- function (x = tmp_stab_indir_table) {
  x <- x[, c("variable", names(x)[!names(x) %in% "variable"])]

  rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
  names(rm_n) <- names(get_sem_var_name_replacement()) 
  x$variable <- str_replace_all(x$variable, rm_n)

  colnames(x) %<>% str_remove_all(., pattern = "via_")

  col_replacement <- c(
    variable = "Variable",
    cv_sp = "CVsp",
    sync = "Synchrony",
    richness = "Horizontal diversity",
    com = "Vertical diversity",
    richness_com = "Horizontal & vertical diversity",
    total = "Total effect"
  )
  colnames(x) %<>% str_replace_all(., col_replacement)

  x
}

kable(format_table(tmp_stab_indir_table),
  format = "latex",
  booktabs = T,
  label = paste0("indir-stab"),
  caption = caption_indir_stab_sem 
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

```{r indir-bm}

kable(format_table(tmp_bm_indir_table),
  format = "latex",
  booktabs = T,
  label = paste0("indir-bm"),
  caption = caption_indir_bm_sem
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

```{r get-sem-single-effect}
stab_sem_coef <- filter(coef_stab_sem, P.Value < 0.05)
get_clean_sem_coef <- function(sem = NULL, resp = NULL, pred = NULL) {
  out <- sem[sem$Response == resp & sem$Predictor == pred, ]$Std.Estimate
  out <- ifelse(length(out) == 0, 0, out)
  round(out, 2)
}

sync_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_sync")
cv_sp_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_cv_sp")
RC1_on_sync <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_sync", pred = "log_RC1")
```

@@

Average trophic level had a positive effect on total biomass ($r_\delta
= `r tlvl_on_bm`$) but connectance had no significant effect (Figure
\@ref(fig:sem-plot) D). Average trophic level had a total positive effect on
temporal stability ($r_\delta = `r com_on_stab$via_cv_sp["t_lvl"]`$, Table
\@ref(tab:indir-stab)) by decreasing $\overline{CV_{sp}}$ ($r_\delta = -0.26$,
Figure \@ref(fig:sem-plot) C). Again, connectance had no significant effect on
temporal stability (Figure \@ref(fig:sem-plot) C, Table SXX).

@@

## Horizontal and vertical diversity had common, opposite and independant effects on total biomass and its temporal stability

@@

We found that connectance covaries negatively with species richness
[@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010],
connectance had itself no effects on both total biomass and its temporal
stability. On the other side, we found species richness was not related to
average trophic level but average trophic level covariated positively with total
biomass and its temporal stability. It indicates that the effects of horizontal
and vertical diversity had independant pathways and the
importance to consider both horizontal and vertical diversity.

>>>>

```{r, echo = FALSE, results = FALSE, message = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

@@

```{r, results = FALSE}
knitr::opts_chunk$set(
  results = TRUE 
)
```

@@

Biodiversity is a key to determine ecosystem properties
[@duffy_why_2009]. Current biodiversity losses and global changes question the
way in which ecosystems will be affected in turn [@bellard_impacts_2012]. Since
the seminal of paper of @may_will_1972, the relationships between
biodiversity and ecosystem properties have received a lot of attention [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018]. Because biodiversity largely
determines both the quantity and the distribution of matter and energy across
ecosystems, it has a tremendous effect on the properties exhibited by an
ecosystem, such as its total biomass [@hector_plant_1999;
@tilman_biodiversity_2014; @grace_integrative_2016;
@pennekamp_biodiversity_2018]], its biomass production
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016]
and the temporal stability of its biomass [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006;
@olivier_independent_nodate].

@@

Biodiversity can be quantified in numerous ways, as the number of species, i.e.
species richness and in number of functional groups, i.e. species are
gathered on what they do [@tilman_biodiversity_2014]. Those metrics can be
gathered into "horizontal diversity", describing the list of components present
in a community, i.e. here species and functional groups. An orthogonal
perspective, based on trophic chains, described species on what they eat. It can
be defined as interactions (or biomass/energy fluxes) between species and how
they are distributed among communities. It can be quantified as the number of
interaction between species or the number of trophic level. This orthogonal view
of seeing biodiversity, can be gathered into "vertical diversity". Horizontal
and vertical diversity are, however, not independent from each other
[@duffy_functional_2007; @zhao_horizontal_2019]. The add or removal of one
species can have different effects depending on its vertical position in the
trophic chain [@duffy_functional_2007]. Hence, it is crucial to study their
interplay to improve our understanding of ecosystem dynamics. 
(*This paragraph is surely too simplistic*)

@@

Environment also largely structure communities and is important to take in
account when considering natural communities. Environment directly affects
organism performance, and subsequently, abundance of species and in turn
determines total biomass [Add general ref; @grace_integrative_2016] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @friedpetersen_drivers_2020]. Hydrology of streams
constraints the composition of fish communities [REF toulouse à balle], and has
already been related to fish population temporal stability [@franssen_annual_2011].
However it is not clear which is the relative strength of direct and indirect
effects of environment on total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

@@

We used an standardized monitoring of fish communities in french streams. 99
sites were followed more than 10 years on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow and enrichment. Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

@@

```{r}
# Basin
st_basin <- get_basin_station(sf_obj = FALSE)

# Select fishing operation 
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)

@@

```{r, fig.show = "hide"}
myload(
  hab_analysis,
  temporal_station_desc,
  temporal_press_polluants,
  geo_station,
  dir = mypath("data")
)
habitat_press <- build_habitat_pressure_dataset(
  .habitat_analysis = hab_analysis,
  .tmp_press = temporal_press_polluants,
  .tmp_st_desc = temporal_station_desc,
  .geo_st = geo_station 
)

# Make PCA 
habitat_press %<>% na.omit()
mask <- names(habitat_press) %in% names(get_pca_var_name_replacement())

pca_rotated <- compute_rotated_pca(
  .data = habitat_press[, mask],
  naxis = 2
)
habitat_pressure <- 
  cbind(habitat_press[, c("station", "temperature_med", "alt")], pca_rotated$rotated$score) %>%
  as_tibble
ggplot(habitat_pressure, aes(x = RC2, y = temperature_med)) +
  geom_point()

ggplot(
  pivot_longer(habitat_pressure, cols = c("RC1", "RC2"),
    names_to = "variable", values_to = "value"
  )
  , aes(x = value)) +
geom_histogram() +
facet_grid(cols = vars(variable))

```{r}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))
tot_var_expl_pca <- sum(prop_var_expl_pca) 
```

```{r, message = FALSE, results = FALSE}
# Com
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb,
 type_network_metrics = "classes"
)
# Compute sem dataset
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync_std"]],
  nmds = NULL,
  basin = st_basin
  )

@@

@@

For each fish station, environmental variables were matched with the year of
fish sampling (except for water temperature for which the record began in 2006).
Then, the average temperature, flow, BOD, river width and depth were computed as
well as coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and
environmental variables are often collinear, a principal component analysis (PCA)
was performed on the scaled variables. We kept in the analysis the two first
principal components based on eigen values distribution (Fig. SXX).
The PCA explained `r tot_var_expl_pca`%
of the variance. An varimax axis rotation was performed to maximize the representation
of the variables by the principal components (Fig. \@ref(fig:pca-rotated)).
Five principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature, (3) average enrichment (BOD) and CV of flow, (4) river size
CV (CV of river width and depth) and (5) CV of enrichment (CV of BOD). Further
informations about data transformation and filtering are detailed in
supplementary information.

@@

Structural Equation Modeling are able to assess multivariate relationships,
directs and indirects between variables.  The metamodel describes how variables
are interrelated (Figure \@ref(fig:matmet) A). We built separated structural
equation model for total biomass and its temporal variability (Figure
\@ref(fig:sem-plot) A & B). Structural equation modeling was set up according to the
hypothesis described in introduction and detailed in supplementary information.
Structural equation models were based on Gaussian linear models, some variables
were logged to fulfill model assumptions (see residual plots and data
transformation in supplementary
informations). To account for heterogeneity of hydrographic basin, we set it as random effect on the intercept. Moreover, statistical model were checked for the presence of
multicollinearity (Supplementary information). The slopes reported in the main
text are standardized in order to compare their magnitude ($r_\delta = \beta
\frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope, $\sigma_x$
and $\sigma_y$ being respectively the standard deviation of the predictive
variable $x$ and of the response variable $y$). In the main text, only the
slopes which had $P$ value inferior to $0.05$ were reported (but see Table SXX).

@@

Environmental variables were interpolated with the `openStars` and `SSN`
packages. Linear models and SEMs were computed with respectively the `nlme` and
`piecewieseSEM` R packages. The code used for the analysis and the Rmarkdown
used to generate this document are accessible on github.

@@

@@

def_env_x_pos <- c(from = 0, to = 3)
asp <- 1.6

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_stab_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "stab_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2")
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_bm_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "bm_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2")
)
```

@@

@@

rich_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_rich_tot_std")
rich_on_tlvl <- get_clean_sem_coef(sem = bm_sem_coef, resp = "t_lvl", pred = "log_rich_tot_std")
tlvl_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "t_lvl")
RC1_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC1")
RC2_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC2")
```

@@

caption_indir_bm_sem <- paste0(
  "Indirect and total effect of variables on total biomass from the
  structural equation model. See the table 1 for details about computation and
  meaning." 
)
```

@@

```{r indir-stab}
kable(format_table(tmp_stab_indir_table),
  format = "latex",
  booktabs = T,
  label = paste0("indir-stab"),
  caption = caption_indir_stab_sem 
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

```{r indir-bm}
kable(format_table(tmp_bm_indir_table),
  format = "latex",
  booktabs = T,
  label = paste0("indir-bm"),
  caption = caption_indir_bm_sem
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

```{r get-sem-single-effect}
stab_sem_coef <- filter(coef_stab_sem, P.Value < 0.05)

sync_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_sync")
cv_sp_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_cv_sp")
RC1_on_sync <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_sync", pred = "log_RC1")
tlvl_on_cv_sp <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_cv_sp", pred = "t_lvl")
```

@@

Average trophic level had a positive effect on total biomass ($r_\delta
= `r tlvl_on_bm`$) but connectance had no significant effect (Figure
\@ref(fig:sem-plot) D). Average trophic level had a total positive effect on
temporal stability ($r_\delta = `r com_on_stab$via_cv_sp["t_lvl"]`$, Table
\@ref(tab:indir-stab)) by decreasing $\overline{CV_{sp}}$
($r_\delta = `r tlvl_on_cv_sp`$,
Figure \@ref(fig:sem-plot) C). Again, connectance had no significant effect on
temporal stability (Figure \@ref(fig:sem-plot) C, Table SXX).

@@

## Horizontal and vertical diversity had common, opposite and dependant effects on total biomass and its temporal stability

@@

We found that connectance covaries negatively with species richness
[@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010],
connectance had itself no effects on both total biomass and its temporal
stability. On the other side, we found species richness was negatively related
to average trophic level and average trophic level covariated positively with
total biomass and its temporal stability. Hence, horizontal diversity had a
direct positive and negative effect respectively on total biomass and its
stability but a negative and positive effect through vertical diversity. It
indicates the importance to consider both horizontal, vertical diversity and
their dependance.
